Perturbed Gait Data

### 2014-10-16 paper.tex
Cleaned up some compile errors/warnings.

% Editing notes
% - No tabs, just spaces!
% - No hanging whitespace
% - Wrap lines to 79 characters
% - Indentation should be two spaces per indent
% - Don't worry about formatting details at first, just get the content in
%   place.

@@

\usepackage{todo}

@@

  Herein we share a data set of typical gait data for 11 subjects walking at a
  three speeds on an instrumented treadmill while being longitudinally
  perturbed in the stance phase with pseudorandom fluctuations in the
  treadmill belt speed. We provide raw marker and ground reaction loads in
  addition to processed data that includes gait landmarks, 2D joint angles,
  angular rates, and joint torques. The protocol is described in detail along
  with the the additional meta data about each of the data files. This data can
  be useful for validating or genreating mathematical models that capable of
  simulating non-periodic and pertubed gaits.

@@

\section{Introduction}
%
Emerging powered prothetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an able-bodied person. A modern gait lab is able to
collect a variety of kinmetic, kinetic, and physiological data from humans
during gait. This data can be used to drive the design of the controller. With
a rich enough data set, one may be able to identify control mechanism used
during a human's natural gait and recovery from perturbations. We have
collected one such data set that we believe is richer than previous gait data
sets and may be rich enough for control identification. The data can also be
used for verifciation purposes for controllers that have been designed in other
manners.

@@

\section{Equipment}
%
The Human Motion and Control Lab at Cleveland State University has such a
system. Our system includes:

\begin{itemize}
  \item A ForceLink R-Mill which has dual 6 DoF force plates, independent belts
    for each foot, and lateral and pitch motion capabilities.
  \item A 10 Camera Motion Analysis motion capture system which includes the
    Cortex software and hardware for collecting analog and camera data
    simultaneously.  Delsys wireless EMG + 3D Accelerometers.
  \item Motek Medical’s D-Flow software and visual display system.
\end{itemize}

Cortex alone is capable of delivering data from the cameras, force plates, and
analog sensors (EMG/Accelerometer), but D-Flow is required to collect data from
digital sensors and the treadmill’s motion (lateral, pitch, and belts). D-Flow
can output multiple types of files which contain the different data.

Our motion capture system’s coordinate system is such that the X coordinate
points to the right, the Y coordinate points upwards, and the Z coordinate
follows from the right-hand-rule, i.e. points backwards with respect to a
person walking forward on the treadmill. The camera’s coordinate system is
aligned to an origin point on treadmill’s surface during camera calibration.

\section{Raw Data}
%
The raw data consists of a set of ASCII tab delimited text files output from
both the Mocap and Record modules in D-Flow in addition to a YAML file that
contains all of the necessary meta data for the given trial. These three files
are stored in a hierarchy of directories with one trial per directory. The
directories are named in the following fashion \verb+T001/+ where \verb+T+
stands for ``trial'' and the following three digits are provide a unique trial
identification number.

@@

@@

@@

@@

\section{Perturbation Signals}

  The data set uses a testing protocol containing a segment of normal 
  walking, followed by a series of pseudorandom longitudinal 
  perturbations, and ending with a second segment of normal walking.   
  Three pseudorandom speed control signals, with mean velocities of 
  0.8, 1.2, and 1.6 m/s, were generated through MATLAB and Simulink. 
  Initially, a random acceleration signal was created using discrete-time 
  Gaussian white noise and saturated at the maximum belt acceleration 
  of 15 m/s2, integrated to obtain belt speed, and high-pass filtered 
  (second-order Butterworth) to eliminate velocity drift.   Mean velocity was 
  then added to the signal and limited between 0 and 3.6 m/s.  The cutoff
  frequencies of the high-pass filter, as well as the variance in the 
  acceleration signal, were adjusted until the appropriate standard deviations
  for each mean velocity were obtained.

  These standard deviations were determined through a trial and error method.
  Initially, perturbations of $\pm$ 10% for each mean velocity were generated. 
  However, this large deviation in the slowest mean velocity (0.8 m/s) would
  unintentionally reduce the treadmill speed to values close to zero.  This 
  would produce an undesired stopping and starting sensation.  Alternatively, 
  smaller standard perturbations of $\pm$ 5% did not produce noticeable 
  perturbations in the fastest trial (1.6 m/s).  Therefore, standard deviations
  of 7.16%, 10.23%, and 13.1% were selected for the 0.8, 1.2, and 
  1.6 m/s speeds, respectively.   This ensured that the test subject was
  sufficiently perturbed for each speed, while remaining within the limits of 
  our testing protocol.

  To ensure that the treadmill could follow the higher frequencies of the 
  random signals, the maximum treadmill acceleration of 15 m/s2 was used in the
  D-Flow Software.  As this value remains constant throughout the experiment,
  the initial starting and stopping of the treadmill would be too rapid for test
  subjects.  To eliminate this, a ramped acceleration and deceleration was 
  generated and placed at the beginning and end of the entire signal,
  respectively.   Other additions include the one minute segments of 
  unperturbed walking at the beginning and end of the trial and 30 seconds of 
  no movement for force plate zeroing and subject T-pose calibration.  The 
  sampling rate of the simulation was set to 100 Hz, corresponding to the 
  D-Flow software's capabilities.

>>>>

% Editing notes
% - No tabs, just spaces!
% - No hanging whitespace
% - Wrap lines to 79 characters
% - Indentation should be two spaces per indent
% - Only one space after periods.
% - Don't worry about formatting details, just get the content in place.

@@

\usepackage{booktabs} % for table /toprule, /midrule, etc
\usepackage{todo}

\title{Perturbed Gait Data}

@@

  Herein we share a typical gait data set collected from 11 subjects walking at
  a three speeds on an instrumented treadmill while being longitudinally
  perturbed during each stance phase with pseudo-random fluctuations in the
  treadmill belt. We provide raw marker and ground reaction loads in addition
  to a presentation of processed data that includes gait landmarks, 2D joint
  angles, angular rates, and joint torques and the software that does the
  processing. The protocol is described in detail along with the additional
  meta data about each of the data files. This data can be useful for
  validating or generating mathematical models that capable of simulating
  non-periodic and perturbed gaits.

@@

\section{Introduction}
%
Emerging powered prosthetics are in need of bio-inspired control systems.
Ideally, a powered prosthetic would behave in such a way that the user would
feel like their limb was never disabled. There are a variety of approaches to
developing bio-inspired control systems some of which aim to mimic the
reactions and motion of an able-bodied person. A modern gait lab is able to
collect a variety of kinematic, kinetic, and physiological data from humans
during gait. This data can potentially be used to drive the design of the human
mimicking controller. With a rich enough data set, one may be able to identify
control mechanism used during a human's natural gait and recovery from
perturbations. We have collected data that we believe is richer than previous
gait data sets and may be rich enough for control identification. The data can
also be used for verification purposes for controllers that have been designed
in other manners.

@@

\section{Equipment}
%
The Human Motion and Control Lab at Cleveland State University has modern human
motion measurement equipment. Our system includes:

\begin{itemize}
  \item A ForceLink R-Mill which has dual 6 degree of freedom force plates,
    independent belts for each foot, and lateral and pitch motion capabilities.
  \item A 10 Camera Motion Analysis motion capture system which includes Osprey
    cameras, the Cortex software, and hardware for collecting analog and camera
    data simultaneously. Delsys wireless EMG + 3D Accelerometers.
  \item Motek Medical’s D-Flow software and visual display system.
\end{itemize}

Cortex alone is capable of delivering data from the cameras along with force
plates, and analog sensors (EMG/Accelerometer) through a National Instruments
XXXX data acquisition unit. D-Flow is required to collect data from any digital
sensors and the treadmill's motion (lateral, pitch, and belts). D-Flow can
output multiple types of files which contain the different data.

Our motion capture system's coordinate system is such that the X coordinate
points to the right, the Y coordinate points upwards, and the Z coordinate
follows from the right-hand-rule, i.e. points backwards with respect to a
person walking forward on the treadmill. The camera's coordinate system is
aligned to an origin point on treadmill's surface during camera calibration.

\section{Raw Data}
%
The raw data consists of a set of ASCII tab delimited text files output from
both the ``mocap'' and ``record'' modules in D-Flow in addition to a YAML file
that contains all of the necessary meta data for the given trial. These three
files are stored in a hierarchy of directories with one trial per directory.
The directories are named in the following fashion \verb+T001/+ where \verb+T+
stands for ``trial'' and the following three digits are provide a unique trial
identification number.

@@

  \small

@@

    \toprule
    Number & Label & Name & Description \\
    \midrule

@@

    \bottomrule

@@

\section{Perturbation Signals}
%
The data set uses a testing protocol containing a segment of normal walking,
followed by a series of pseudo-random longitudinal perturbations, and ending
with a second segment of normal walking. Three pseudo-random speed control
signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
generated through MATLAB and Simulink. Initially, a random acceleration signal
was created using discrete-time Gaussian white noise and saturated at the
maximum belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt
speed, and high-pass filtered (second-order Butterworth) to eliminate velocity
drift. Mean velocity was then added to the signal and limited between 0 and 3.6
m/s. The cutoff frequencies of the high-pass filter, as well as the variance in
the acceleration signal, were adjusted until the appropriate standard
deviations for each mean velocity were obtained.

These standard deviations were determined through a trial and error method.
Initially, perturbations of $\pm$ 10\% for each mean velocity were generated.
However, this large deviation in the slowest mean velocity (0.8 $m \cdot
s^{-1}$) would unintentionally reduce the treadmill speed to values close to
zero. This would produce an undesired stopping and starting sensation.
Alternatively, smaller standard perturbations of $\pm$ 5\% did not produce
noticeable perturbations in the fastest trial (1.6 $m \cdot s^{-1}$).
Therefore, standard deviations of 7.16\%, 10.23\%, and 13.1\% were selected for
the 0.8, 1.2, and 1.6 $m \cdot s^{-1}$ speeds, respectively.  This ensured that
the test subject was sufficiently perturbed for each speed, while remaining
within the limits of our testing protocol.

To ensure that the treadmill could follow the higher frequencies of the random
signals, the maximum treadmill acceleration of 15 $m \cdot s^{-2}$ was used in
the D-Flow Software. As this value remains constant throughout the experiment,
the initial starting and stopping of the treadmill would be too rapid for test
subjects. To eliminate this, a ramped acceleration and deceleration was
generated and placed at the beginning and end of the entire signal,
respectively. Other additions include the one minute segments of unperturbed
walking at the beginning and end of the trial and 30 seconds of no movement for
force plate zeroing and subject T-pose calibration.  The sampling rate of the
simulation was set to 100 Hz, corresponding to the D-Flow software's
capabilities.
